---
title: 'Simple R Functions '
author: 'Yue Wu'
date: "Feburary 1, 2018"
output: 
  pdf_document: 
    latex_engine: xelatex
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### 1.

(a) Write functions \texttt{tmpFn1} and \texttt{tmpFn2} such that if \texttt{xVec} is the vector $(x_1, x_2, ... ,x_n)$, then \texttt{tmpFn1(xVec)} returns vector $(x_1, x_2^2, ..., x_n^n)$ and \texttt{tmpFn2(xVec)} returns the vector $(x_1, \frac{x_2^2}{2}, ..., \frac{x_n^n}{n})$.

******

Here is \texttt{tmpFn1}
```{r}

tmpFn1 <- function(xVec){
  return(xVec^(1:length(xVec)))
}

## simple example
a <- c(2, 5, 3, 8, 2, 4)

b <- tmpFn1(a)
b

```

and now \texttt{tmpFn2}

```{r}
tmpFn2 <- function(xVec2){
  
  n = length(xVec2)
  
  return(xVec2^(1:n)/(1:n))
}



c <- tmpFn2(a)
c


```

######  
  
  
  
(b) Now write a fuction \texttt{tmpFn3} which takes 2 arguments $x$ and $n$ where $x$ is a single number and $n$ is a strictly positive integer.  The function should return the value of   
$$1 + \frac{x}{1} + \frac{x^2}{2} + \frac{x^3}{3} + ... + \frac{x^n}{n}$$
******

Here is \texttt{tmpFn3}
```{r}

tmpFn3 <- function(x, n){
  return(sum(1+x^(1:n)/(1:n)))
}

## simple example
b <- tmpFn3(5,5)
b

```



\newpage
### 2. Write a funciton \texttt{tmpFn(xVec)} such that if \texttt{xVec} is the vector $x = (x_1, ...,x_n)$ then \texttt{tmpFn(xVec)} returns the vector of moving averages:

$$\frac{x_1 + x_2 + x_3}{3}, \frac{x_2 + x_3 + x_4}{3}, ...,\frac{x_{n-2} + x_{n-1} + x_n}{3}$$
******

Here is \texttt{tmpFn}
```{r}

tmpFn <- function(xVec){
  l = length(xVec)
  return(sum((xVec[1:(l-2)]+xVec[2:(l-1)]+xVec[3:l])/3))
}

## example
b <- tmpFn(c(1:5,6:1))
b

```


### 3. Consider the continuous function

$$f(x) = \left\{\begin{matrix}
x^2 + 2x + 3 & if  & x < 0 \\ 
x + 3 & if & 0 \leq x < 2 \\ 
x^2 + 4x - 7 & if & 2 \leq  x 
\end{matrix}\right.$$

Write a function \texttt{tmpFn} which takes a single argument \texttt{xVec}.  the function should return the vector the values of the function $f(x)$ evaluated at the values in \texttt{xVec}.  
Hence plot the function $f(x)$ for $-3 < x <3$.
******

Here is \texttt{tmpFn}
```{r}

tmpFn <- function(xVec){
  
  if (xVec<0) {
    return(xVec^2+2*xVec+3)
  }
  else if ((xVec>=0) & (xVec<2)){
    return (xVec+3)
  }
  else if (xVec>=2) {
    return (xVec^2+4*xVec-7)
  }
}

## example
tmp <- seq(-3, 3, len=30)
plot(tmp, tmpFn(tmp), type="l")

```

### 4. Write a function which takes a single argument which is a matrix.  The function should return a matrix which is the same as the function argument but every odd number is doubled.  
Hence the result of using the function on the matrix   

$$\begin{bmatrix}
1 & 1 & 3\\ 
5 & 2 & 6\\ 
-2 & -1 & -3
\end{bmatrix}$$

should be:

$$\begin{bmatrix}
2 & 2 & 6\\ 
10 & 2 & 6\\ 
-2 & -2 & -6
\end{bmatrix}$$
******

Here is \texttt{dodd}
```{r}

dodd <- function(mat){
  mat[mat%%2==1] <- mat[mat%%2==1]*2
  return(mat)
}

## example
tmp <- matrix(c(1,1,3,5,2,6,-2,-1,-3),ncol=3,nrow=3,byrow=TRUE)
dodd(tmp)

```

### 5. Write a function which takes 2 arguements $n$ and $k$ which are positive integers.  It should return the $\mathit{n} x \mathit{n}$ matrix:


$$\begin{bmatrix}
k & 1 & 0 & 0 & \cdots  & 0 & 0\\ 
1 & k & 1 & 0 & \cdots  & 0 & 0\\
0 & 1 & k & 1 & \cdots  & 0 & 0\\
0 & 0 & 1 & k & \cdots  & 0 & 0\\
\cdot & \cdot & \cdot & \cdot & \cdot & \cdot & \cdot\\ 
0 & 0 & 0 & 0 & \cdots  & k & 1\\
0 & 0 & 0 & 0 & \cdots  & 1 & k
\end{bmatrix}$$
******

Here is \texttt{mat}
```{r}

mat <- function(n,k){
  m <- matrix(0,ncol=n,nrow=n)
  m[col(m)==row(m)] <- k
  m[col(m)==row(m)+1] <- 1
  m[col(m)==row(m)-1] <- 1
  return(m)
}

## example
tmp <- mat(5,3)
tmp

```

### 6. Suppose an angle $\alpha$ is given as a positive real number of degrees.  
If $0 \leq \alpha < 90$ then it is quadrant 1.  If $90 \leq \alpha < 180$ then it is quadrant 2.  
if $180 \leq \alpha < 270$ then it is quadrant3.  if $270 \leq \alpha < 360$ then it is quadrant 4.  
if $360 \leq \alpha < 450$ then it is quadrant 1.  
And so on ...   

Write a function \texttt{quadrant(alpha)} which returns the quadrant of the angle $\alpha$.
******

Here is \texttt{quad}
```{r}

quad <- function(a){
  if (((a+90)%/%90)%%4 ==0) {
    return(4)
  }
  else{
    return(((a+90)%/%90)%%4)
  }
}
## example
tmp <- quad(271)
tmp

```

### 7.

(a) Zeller's congruence is the formula:

$$ f = ([2.6m - 0.2] + k + y + [y/4] + [c/4] - 2c)mod 7 $$
where $[x]$ denotes the integer part of $x$; for example $[7.5] = 7$.  

Zeller's congruence returns the day of the week $f$ given:

$k$ = the day of the month  
$y$ = the year in the century  
$c$ = the first 2 digits of the year (the century number)  
$m$ = the month number (where January is month 11 of the preceding year, February is month 12 of the preceding year, March is month 1, etc.)   
For example, the date 21/07/1`963 has $m = 5, k = 21, c = 19, y = 63$;  
the date 21/2/63 has $m=12, k=21, c=19, and y=62$.  
Write a function \texttt{weekday(day,month,year)} which returns the day of the week when given the numerical inputs of the day, month and year.  
Note that the value of 1 for $f$ denotes Sunday, 2 denotes Monday, etc.
(b) Does your function work if the input parameters \text{day, month, and year} are vectors with the same length and valid entries?


******

Here is \texttt{weekday}
```{r}

weekday <- function(day,month,year){
  month <- month - 2
  year[month<=0] <- year-1
  month[month<=0] <- month+12
  c <- year %/% 100
  y <- year - 100*c
  d <- ((2.6*month-0.2)%/%1+day+y+y%/%4+c%/%4-2*c)%%7
  return(c("Sunday","Monday","Tuesday","Wednesday","Thursday","Friday","Saturday")[1+d])
}
## example
weekday(21,07,1963)

## b: with vector input
weekday(c(21,07,1963),c(22,07,1963),c(31,01,2018))

```

### 8.  (a) Suppose $x_0 = 1$ and $x_1 = 2$ and  
$$x_j=x_{j-1} + \frac{2}{x_{j-1}}$$
Write a function testLoop which takes the single argument n and returns the first $n − 1$ values of the
sequence ${x_j}_j≥0$: that means the values of $x_0, x_1, x_2, . . . , x_{n−2}$.
## (b) Now write a function testLoop2 which takes a single argument yVec which is a vector. The function should return
$$\sum_{j=1}^{n} e^j$$
where $n$ is the length of \texttt{yVec}.

******

Here is \texttt{testLoop}
```{r}
# (a)
testLoop <- function(n){
  v <- rep(NA, n-1)
  v[1] <- 1
  v[2] <- 2
  for( j in 3:(n-1) )
    v[j]= v[j-1]+2/v[j-1]
  return(v)
  }
# test example
testLoop(8)

# (b)
testLoop2 <- function(yVec){
  n= length(yVec)
  v <- c(1:n)
  v[1:n] <- exp(v[1:n])
  return(sum(v))
  }
# test example
testLoop2(c(1:8))
```

### 9. Solution of the difference equation $x_n = rx_{n−1}(1 − x_{n−1})$, with starting value $x_1.$
##(a) Write a function \texttt{quadmap( start, rho, niter )} which returns the vector $(x_1, . . . , x_n)$ where $xk = rx_{k−1}(1 − x_{k−1})$ and
  \texttt{niter} denotes $n$,
  \texttt{start} denotes $x_1$, and
  \texttt{rho} denotes $r$.
  Try out the function you have written:
  • for $r = 2$ and $0 < x1 < 1$ you should get $x_n → 0.5$ as $n → {\infty}$.
  • try \texttt{tmp <- quadmap(start=0.95, rho=2.99, niter=500)}
  Now switch back to the Commands window and type:
  \texttt{plot(tmp, type="l")}
  Also try the plot $plot(tmp[300:500], type="l")$
##(b) Now write a function which determines the number of iterations needed to get $| x_n − x_{n−1} |< 0.02$. So this function has only 2 arguments: $start$ and $rho$. $(For start=0.95$ and  $rho=2.99$, the answer is $84.)$
******

Here is \texttt{quadmap}
```{r}
# (a)
quadmap <- function(start, rho, niter)
  {
  v <- rep(NA,niter)
  v[1] <- start
  for(i in 1:(niter-1)) {
  v[i + 1] <- rho*v[i]*(1-v[i])
  }
  return(v)
  }
# test example
quadmap(0.5,2,100)[90:100]
tmp <- quadmap(start=0.95, rho=2.99, niter=500)
plot(tmp, type="l")
plot(tmp[300:500], type="l")

# (b)
quadmap2 <- function(start, rho)
  {
  x1 <- start
  niter <- 1
  while(abs(x1-rho*x1*(1- x1))>= 0.02) {
  x1 <- rho*x1*(1- x1)
  niter <- niter+ 1
  }
  return(niter)
}
# test example
quadmap2(0.95,2.99)
```

### 10. (a) Given a vector (x1, . . . , xn), the sample autocorrelation of lag k is defined to be
$$r_k = \frac{\sum_{i=k+1}^{n}(x_i-\bar{x})(x_{i-k}-\bar{x})}{\sum_{i=1}^{n}(x_i-\bar{x})^2}$$
  Thus
$$r_1 = \frac{\sum_{i=2}^{n}(x_i-\bar{x})(x_{i-1}-\bar{x})}{\sum_{i=1}^{n}(x_i-\bar{x})^2} = \frac{(x_2-\bar{x})(x_1-\bar{x})+...+(x_n-\bar{x})(x_n-1-\bar{x})}{\sum_{i=1}^{n}(x_i-\bar{x})^2}$$
  Write a function \texttt{tmpFn(xVec)} which takes a single argument \texttt{xVec} which is a vector and returns a \texttt{list} of two values: $r_1$ and $r_2$. In particular, find $r_1$ and $r_2$ for the vector $(2, 5, 8, . . . , 53, 56)$
##(b) (Harder.) Generalise the function so that it takes two arguments: the vector \texttt{xVec} and an integer \texttt{k} which lies between $1$ and $n − 1$ where n is the length of xVec. 
The function should return a vector of the values $(r_0 = 1, r_1, . . . , r_k)$.
If you used a loop to answer part (b), then you need to be aware that much, much better solutions are
possible—see exercises 4. (Hint: \texttt{sapply}.)
******

Here is \texttt{tmpFn}
```{r}
# (a)
tmpFn <- function(xVec)
  {
  xbar <- xVec - mean(xVec)
  xbar2 <- sum(xbar^2)
  n <- length(xVec)
  r1 <- sum( xbar[2:n] * xbar[1:(n-1)] )/xbar2
  r2 <- sum( xbar[3:n] * xbar[1:(n-2)] )/xbar2
  return(list(r1 = r1, r2 = r2))
  }
# test example
tmpFn(seq(2,56,by=3))

# (b)
tmpFn <- function(x,k)
  {
  xbar <-  x - mean(x)
  xbar2 <- sum(xbar^2)
  n <- length(x)
  return(c(1,sapply(1:k, function(y) sum(xbar[(y+1):n]*xbar[1:(n-y)] )/xbar2)))
  }
# test example
tmpFn(seq(2,56,by=3),3)
```




